The Private School Network: Recruiting Visits to Private High Schools by Public and Private Universities

### 2023-12-15 r_r/manuscript_rihe_r_r.Rmd
updates to manuscript

# Data and sample

@@

The data collection sample for the broader project consisted of: all public research-extensive universities as defined by the 2000 Carnegie Classification (N=102), all private universities in the top 100 of U.S. News and World Report National Universities rankings (N=58), and all private colleges in the top 50 of U.S. News and World Report Liberal Arts Colleges rankings (N=47).<!--TODO: Confirm N's--> The rationale for different criteria for inclusion across different university types is because these criteria reflected the interest of a foundation that funded the data collection. This article excludes private liberal arts colleges because having a third institutional type would substantially increase article length. The final analysis sample for this article consists of 15 public research universities and 14 private research universities that we believed posted all varieties of off-campus recruiting events on their admissions website. For these universities, we analyze recruiting visits to private high schools that meet the following criteria: enrolls at least 10 12th grade students, is not primarily a virtual school, and is located in the 50 U.S. states, the District of Columbia, or land regulated by the Bureau of Indian Affairs.

@@

This article faces several limitations. First, our data collection sample consisted of public research-extensive universities as defined by the 2000 Carnegie Classification (N=102) and all private universities in the top 100 of U.S. News and World Report National Universities rankings (N=58). Making comparisons between public and private universities would be more appropriate if we used the same classification method to define the data collection sample for all universities. The process we used to define the data collection sample yielded a sample of public research universities with lower USNWR rankings than the analysis sample of selective private universities. However, the columns in Table \@ref(tab:sample-characteristics) for 25th and 75th percentile SAT score of enrolled freshmen suggest that private universities ranked lower than 75 have test score ranges that are fairly similar to public research universities not ranked in the top 100.

@@

INSERT FIGURES \@ref(fig:race-comp-pubu-privhs-pubhs) AND \@ref(fig:race-comp-privu-privhs-pubhs) ABOUT HERE

@@

Research question 1 asks about the scale of visits to private high schools. All private universities in our sample made a disproportionate number of visits to private high schools (Figure \@ref(fig:actual-proportional-privu)) and, for about half of these universities, visits to private high schools represented about 50% or more of all high school visits. For most public research universities in our sample, the scale and relative scale -- as a percentage of all HS visits -- of visits to private high schools were smaller than that of most selective private universities. However, 13 of the 15 public universities visited a disproportionate number of private high schools (Figure \@ref(fig:actual-proportional-pubu)) and that vast majority of these private school visits were to out-of-state private schools. Additionally, several public research universities -- University of Alabama, CU-Boulder, University of South Carolina -- visited private high schools at a scale that exceeded most private universities.

Research question 2 asks about overlap in the private school recruiting networks of public and private universities. We observed substantial overlap and variation in recruiting networks of public universities. For example, the recruiting networks of the University of Arkansas and the University of Kansas share substantial overlap and are very different from the recruiting networks of UMass-Amherst and Rutgers, which share substantial overlap with one another. Amongst the network of public and private universities, we observe substantial overlap and variation in recruiting networks. Figure \@ref(fig:plot-1mode-u) shows that three clusters emerge -- each containing some mix of public and private universities. The universities within a cluster have some overlap in the kinds of schools visited (e.g., TCU and University of South Carolina), but also share substantial overlap with particular universities in different clusters (e.g., Tufts and University of Pittsburgh). Whether we are categorizing universities by cluster analysis or by proximity on the network graph, we observe several different subsets of universities. Each of these subsets is associated with a different approach to private school visits and -- with the exception of a few universities like UC-Riverside and SUNY Stony Brook that don't play the private school game -- each of these subsets contained both public and private universities.

@@

With the exception of UC-Berkeley and UC-San Diego, the academic rating of private schools visited by public universities was lower than the rating of those visited by most private universities. This finding is largely explained by our analysis sample, which included public research universities regardless of rank but only included private universities ranked as a top 100 National University by U.S. News. Nevertheless, the large number of visits to low-rated private schools by public universities, combined with the fact that the overwhelming majority of these visits were to out-of-state schools, is consistent with the idea that many public flagship universities visit private schools primarily in search of full-pay students rather than academic superstars.

Across the universe of private high schools, most schools enroll relatively few students who identify as Black, Latinx, or Native. For the public and private universities in our sample, the set of visited private schools enrolled fewer Black, Latinx, and Native students than the national universe of private schools. By contrast, the public high schools visited by public universities tended to enroll a much higher share of students who identify as Black, Latinx, or Native. [ADD A SENTENCE ABOUT RACIAL COMPOSITION OF VISITS TO SOUTH?]

@@

This study engages with two, previously disconnected literatures. Scholarship on nonresident enrollment observes that many public research universities began growing nonresident enrollment following declines in state appropriations [@RN3753]. Given that nonresident tuition price is often two- to three-times resident price, it is unsurprising that nonresident students tend to be more affluent than resident students, are less likely to be first-generation students, and are less likely to be Black or Latinx [@RN3685]. Public research universities recruit nonresident students using institutional financial aid packages [@RN3762] and through recruiting interventions that target earlier stages in the enrollment funnel. Using nearly the same sample of 15 public research universities used in this study, @RN4758 found that 12 universities made more out-of-state visits than in-state visits and visits to out-of-state public high schools focused on affluent, predominantly white schools. Public research universities have a financial incentive to target private school students, who tend to come from affluent households able to pay nonresident tuition price without need-based aid. This study extends the literature on nonresident students by analyzing the scale of visits to private schools, the overlap and divergence in recruiting networks between public and private universities, and the characteristics of visited private schools.

We also engage with scholarship on linkages between private schools and private universities, a linkage that sociologists have observed for many decades. Private school students comprise a disproportionate share of enrollment at selective private universities (Author calculations based on HSLS:2009). Practically, these linkages exist because private schools want to send their students to "good" universities while selective private universities want to enroll students who can pay private university tuition and who have desired academic profile and/or extracurricular activities [@RN3519]. Prior research has often focused on linkages between elite private schools and elite universities [@RN4407; @RN4935; @RN4640]. Whereas elite universities (e.g., Princeton, Stanford) experience strong student demand based on reputation, merely-selective private universities must incite demand by expending resources on recruiting. Prior research has not systematically analyzed connections between private schools and private universities that are selective, but not elite. Therefore, this study extends the literature by investigating private school visits for a sample of 14 selective private universities that exhibit variation in geographic location, religious affiliation, and ranking.

@@

Our analyses shed light about the nature of this competition. When making decisions about which private schools to visit, public research universities seem to be mimicking the behavior of particular private universities rather than mimicking the behavior of private universities writ large. Several public universities seem to be competing directly for students targeted by a particular private university. For example, UC-Berkeley visited 27.9% of the schools visited by Northwestern, 28.9% of the schools visited by Tufts, and 27.8% of the schools visited by Emory (Figure \@ref(tab:privu-percent-visited-by-pubu-matrix)). We reason UC-Berkeley is competing directly for students because all four universities are in the same broad echelon of prestige. Other potential examples of direct competition include the University of Georgia targeting schools visited by Emory, SMU, TCU, and Tulane, or the University of Arkansas targeting schools visited by Baylor.

However, public universities that made the greatest number of private school visits seemed to be targeting lower-achieving students at private schools targeted by particular private universities. For example, the University of Alabama visited 52.2% of the 625 private schools that Notre Dame visited, with most of the overlap at Catholic schools. Given the differences in academic prestige between them, the University of Alabama may not be competing for the same students as Notre Dame. Instead, they seem to be targeting Catholic school students of a similar economic echelon, who have lower academic achievement, and want to attend a football powerhouse. Additionally, CU-Boulder visited 42.4% of schools visited by Northwestern, which could be interpreted as CU-Boulder competing for lower-achieving students of a similar economic and cultural echelon at the schools targeted by Northwestern. Similarly, the University of South Carolina competes heavily at schools targeted by Emory, SMU, TCU, and Tulane.

@@

Market research suggests that off-campus recruiting visits will remain a vital means of connecting to students and guidance counselors, just as they were prior to the Covid pandemic. @RN4664 found that "travel" was the top-ranked top-ranked marketing/recruiting expenditure item for both public universities and private universities. During the height of Covid, off-campus recruiting visits ceased and were partially replaced by virtual counselor events [@parents_covid_rnl]. Admissions counselors are now traveling again. @RN4896 found that travel was the fifth ranked expenditure item for public universities and the second ranked expenditure item for private universities. When asked about areas of future expenditure growth, travel was the top ranked item for public universities and the second ranked item for private universities [@RN4896], suggesting that universities are increasing travel following the decline of the pandemic.

The @eab_nacac_counselor_2023 survey of high school counselors asked "how do you prefer to receive information from colleges?" High school visits by college representatives was by far the highest rated "channel" for private school counselors and was also the highest rated channel for public school counselors. By contrast, virtual visits by admissions representatives received low ratings by high school guidance counselors. Finally, @RN4895 asked universities to rank the most effective "outreach strategies for high school counselors." "Calling after visits" was deemed effective for 97% of private universities (the top strategy) and was deemed effective by 88% of public universities. What insights do we glean from these market research findings? In the era of mass marketing and data science, the pursuit of traditional college students from privileged schools and communities is still a hand-to-hand business and off-campus recruiting visits remain a vital means of recruiting these students.

Considering the practical importance of off-campus recruiting visits, what are the practical implications of our research? Prior scholarly and market research shows that students are sensitive to recruiting interventions, including recruiting visits [@RN4324; @RN4402]. Consequently, racial and socioeconomic composition of the schools that a university chooses to visit affects the racial and socioeconomic composition of the university. Our research on visits to public high schools [@RN4758] and this study on visits to private schools finds that our sample of public research universities and selective private universities tend to focus their recruiting efforts on affluent and predominantly white schools and communities. Some universities may publicly proclaim a commitment to social mobility and privately pursue an intentional recruiting strategy that targets affluent schools and communities. We offer practical guidance for universities that may be unintentionally implementing a recruiting strategy that privileges the privileged.

Several studies suggest that universities -- particularly elite private universities -- pursue racial diversity by recruiting non-white students who attend predominantly white private high schools [e.g., @RN4495]. Additionally, scholarship that engages with enrollment in private schools by Black students suggests that many high-income Black families began sending their children to elite, predominantly white private schools once these schools were willing to enroll Black students [@RN4724]. Therefore, universities may visit predominantly white private high schools in search of non-white students. One concern about this practice is that non-white students at elite private schools may be in high demand by many universities, but similarly achieving non-white students at predominantly non-white public schools might be ignored

In the aftermath of the Supreme Court decision on race conscious admissions, @RN4954 published a report titled *Strategies for Increasing Diversity and Opportunity in Higher Education.*  The report recommended that universities diversify the set of high schools they recruit from:

>> Rather than recruiting primarily from private and public high schools composed of predominantly high-income students, institutions can expand and prioritize their outreach to high schools with substantial populations of low-income students and students of color. As part of expanding outreach, institutions can assign admissions recruiters to high schools from which they have not typically recruited to provide equal opportunity for all students

@@

kable(univ_sample, caption = "Characteristics of universities in analysis sample", booktabs = T, escape = F, align = 'c') %>%
  collapse_rows(columns = 1, latex_hline = 'major', valign = 'middle') %>%
  column_spec(c(2), width = '4.5cm', latex_valign = 'm') %>% 
  column_spec(c(3), width = '1.8cm', latex_valign = 'm') %>% 
  column_spec(c(4, 5), width = '3.8cm', latex_valign = 'm') %>% 
  column_spec(c(6, 8, 10:14), width = '2.4cm', latex_valign = 'm') %>% 
  column_spec(c(7), width = '2.7cm', latex_valign = 'm') %>% 
  column_spec(c(9, 15), width = '3cm', latex_valign = 'm') %>% 
  kable_styling(font_size = 13, latex_options = c('scale_down', 'hold_position'))
```

@@

# 49 public research
pop_pubu <- c('151351', '171100', '180461', '188030', '199193', '147703', '204857', '207388', '214777', '186380', '149222', '196088', '196097', '204796', '100751', '197036', '164155', '104179', '106397', '110635', '110644', '110653', '110662', '110671', '110680', '201885', '126614', '129020', '139959', '141574', '225511', '142285', '155317', '161253', '163286', '166629', '170976', '181464', '187985', '199120', '215293', '217484', '218663', '230764', '234076', '236948', '230728', '234030', '172644')

# 49 private national
pop_privu <- c('194824', '139658', '190150', '223232', '212054', '193900', '166027', '144050', '130794', '243744', '166683', '198419', '215062', '162928', '182670', '110404', '147767', '217156', '190415', '227757', '152080', '186131', '221999', '131496', '164988', '160755', '213543', '135726', '216597', '228246', '191241', '186867', '131159', '228875', '239105', '127060', '122436', '207971', '179159', '168148', '164924', '195030', '179867', '201645', '197708', '199847', '211440', '121150', '123961')

@@

kable(univ_pop_df, caption = "Characteristics of universities in analysis sample compared to data collection sample", booktabs = T, escape = F, align = 'c', linesep = '') %>%
  add_header_above(header = c(' ' = 1, 'Analysis sample (N=15)' = 3, 'Data collection sample (N=49)' = 3, 'Analysis sample (N=14)' = 3, 'Data collection sample (N=49)' = 3)) %>%
  add_header_above(header = c(' ' = 1, 'Public research universities' = 6, 'Selective private universities' = 6)) %>%
  kable_styling(font_size = 13, latex_options = 'scale_down') 
```

@@

kable(table_1mode_pubu$onemode_pct_df, caption = 'One-mode public university percent matrix', booktabs = T, align = 'c', linesep = '') %>%
  column_spec(c(7, 9, 15), width = '5em') %>%
  column_spec(c(4, 12), width = '5.5em') %>%
  column_spec(c(6, 10, 16), width = '6em') %>%
  column_spec(c(2, 11, 13), width = '6.5em') %>%
  column_spec(c(3, 8), width = '7em') %>%
  column_spec(c(5), width = '7.5em') %>%
  column_spec(c(14), width = '8.8em') %>%
  kable_styling(font_size = 13, latex_options = 'scale_down') 
```

@@

kable(table_1mode_privu$onemode_pct_df, caption = 'One-mode private university percent matrix', booktabs = T, align = 'c', linesep = '') %>%
  column_spec(c(2, 3, 7:9, 11), width = '7em') %>% 
  column_spec(c(4:6, 10, 12, 13), width = '6em') %>% 
  column_spec(14:15, width = '8.5em') %>%
  kable_styling(font_size = 13, latex_options = 'scale_down') 
```

@@

kable(table_1mode_u$onemode_pct_df[c(1,7,9,13,16,17,18,21,22,23,24,25,27,28,29),c(2,3,4,5,6,8,10,11,12,14,15,19,20,26)], caption = 'For private universities, the percent of high schools visited by public universities', booktabs = T, align = 'c', linesep = '') %>%
  column_spec(c(2, 3, 7:9, 11), width = '7em') %>% 
  column_spec(c(4:6, 10, 12, 13), width = '6em') %>% 
  column_spec(14:15, width = '8.5em') %>%
  kable_styling(font_size = 13, latex_options = 'scale_down')
```

@@

kable(table_1mode_u$onemode_pct_df[c(2,3,4,5,6,8,10,11,12,14,15,19,20,26),c(1,7,9,13,16,17,18,21,22,23,24,25,27,28,29)], caption = 'For public universities, the percent of high schools visited by private universities', booktabs = T, align = 'c', linesep = '') %>%
  column_spec(c(7, 9, 15), width = '5em') %>%
  column_spec(c(4, 12), width = '5.5em') %>%
  column_spec(c(6, 10, 16), width = '6em') %>%
  column_spec(c(2, 11, 13), width = '6.5em') %>%
  column_spec(c(3, 8), width = '7em') %>%
  column_spec(c(5), width = '7.5em') %>%
  column_spec(c(14), width = '8.8em') %>%
  kable_styling(font_size = 13, latex_options = 'scale_down')
```

@@

kable(table_1mode_pubu$onemode_df, caption = 'One-mode public university count matrix', booktabs = T, align = 'c', linesep = '') %>%
  column_spec(c(7, 9, 15), width = '5em') %>%
  column_spec(c(4, 12), width = '5.5em') %>%
  column_spec(c(6, 10, 16), width = '6em') %>%
  column_spec(c(2, 11, 13), width = '6.5em') %>%
  column_spec(c(3, 8), width = '7em') %>%
  column_spec(c(5), width = '7.5em') %>%
  column_spec(c(14), width = '8.8em') %>%
  kable_styling(font_size = 13, latex_options = 'scale_down') 
```

@@

kable(table_1mode_privu$onemode_df, caption = 'One-mode private universities count matrix', booktabs = T, align = 'c', linesep = '') %>%
  column_spec(c(2, 3, 7:9, 11), width = '7em') %>% 
  column_spec(c(4:6, 10, 12, 13), width = '6em') %>% 
  column_spec(14:15, width = '8.5em') %>%
  kable_styling(font_size = 13, latex_options = 'scale_down') 
```

>>>>

## Data and Sample

@@

The data collection sample for the broader project consisted of: all public research-extensive universities as defined by the 2000 Carnegie Classification (N=102), all private universities in the top 100 of U.S. News and World Report National Universities rankings (N=57), and all private colleges in the top 50 of U.S. News and World Report Liberal Arts Colleges rankings (N=47). The rationale for different criteria for inclusion across different university types is because these criteria reflected the interest of a foundation that funded the data collection. This article excludes private liberal arts colleges because having a third institutional type would substantially increase article length. The final analysis sample for this article consists of 15 public research universities and 14 private research universities that we believed posted all varieties of off-campus recruiting events on their admissions website. For these universities, we analyze recruiting visits to private high schools that meet the following criteria: enrolls at least 10 12th grade students, is not primarily a virtual school, and is located in the 50 U.S. states, the District of Columbia, or land regulated by the Bureau of Indian Affairs.

@@

This article faces several limitations. First, our data collection sample consisted of public research-extensive universities as defined by the 2000 Carnegie Classification (N=102) and all private universities in the top 100 of U.S. News and World Report National Universities rankings (N=57). Making comparisons between public and private universities would be more appropriate if we used the same classification method to define the data collection sample for all universities. The process we used to define the data collection sample yielded a sample of public research universities with lower USNWR rankings than the analysis sample of selective private universities. However, the columns in Table \@ref(tab:sample-characteristics) for 25th and 75th percentile SAT score of enrolled freshmen suggest that private universities ranked lower than 75 have test score ranges that are fairly similar to public research universities not ranked in the top 100.

@@

<!--TODO: Supplemental figures \@ref(fig:race-comp-pubu-privhs-pubhs) AND \@ref(fig:race-comp-privu-privhs-pubhs) are now moved to appenidix-->

@@

Research question 1 asks about the scale of visits to private high schools. All private universities in our sample made a disproportionate number of visits to private high schools (Figure \@ref(fig:actual-proportional-privu)) and, for about half of these universities, visits to private high schools represented about 50% or more of all high school visits. For most public research universities in our sample, the scale and relative scale -- as a percentage of all HS visits -- of visits to private high schools were smaller than that of most selective private universities. However, 13 of the 15 public universities visited a disproportionate number of private high schools (Figure \@ref(fig:actual-proportional-pubu)) and the vast majority of these private school visits were to out-of-state private schools. Additionally, several public research universities -- University of Alabama, CU-Boulder, University of South Carolina -- visited private high schools at a scale that exceeded most private universities.

Research question 2 asks about overlap in the private school recruiting networks of public and private universities. We observed substantial overlap and variation in the recruiting networks of public universities. For example, the recruiting networks of the University of Arkansas and the University of Kansas share substantial overlap and are very different from the recruiting networks of UMass-Amherst and Rutgers, which share substantial overlap with one another. Amongst the network of public and private universities, we observe substantial overlap and variation in recruiting networks. Figure \@ref(fig:plot-1mode-u) shows that three clusters emerge -- each containing some mix of public and private universities. The universities within a cluster have some overlap in the kinds of schools visited (e.g., TCU and University of South Carolina), but also share substantial overlap with particular universities in different clusters (e.g., Tufts and University of Pittsburgh). Whether we are categorizing universities by cluster analysis or by proximity on the network graph, we observe several different subsets of universities. Each of these subsets is associated with a different approach to private school visits and -- with the exception of a few universities like UC-Riverside and SUNY Stony Brook that don't play the private school game -- each of these subsets contained both public and private universities.

@@

With the exception of UC-Berkeley and UC-San Diego, the average academic rating of private schools visited by public universities was lower than the rating of those visited by most private universities. This finding is largely explained by our analysis sample, which included public research universities regardless of rank but only included private universities ranked in the top 100 National University by U.S. News. Nevertheless, the large number of visits to low-rated private schools by public universities, combined with the fact that the overwhelming majority of these visits were to out-of-state schools, is consistent with the idea that many public flagship universities visit private schools primarily in search of full-pay students rather than academic superstars.

Across the universe of private high schools, most schools enroll relatively few students who identify as Black, Latinx, or Native. For the public and private universities in our sample, the set of visited private schools enrolled fewer Black, Latinx, and Native students than the national universe of private schools. By contrast, the public high schools visited by public universities tended to enroll a much higher share of students who identify as Black, Latinx, or Native. <!--TODO: [ADD A SENTENCE ABOUT RACIAL COMPOSITION OF VISITS TO SOUTH?]-->

@@

This study engages with two, previously disconnected literatures. Scholarship on nonresident enrollment observes that many public research universities began growing nonresident enrollment following declines in state appropriations [@RN3753]. Given that nonresident tuition price is often two- to three-times resident price, it is unsurprising that nonresident students tend to be more affluent than resident students, are less likely to be first-generation students, and are less likely to be Black or Latinx [@RN3685]. Public research universities recruit nonresident students using institutional financial aid packages [@RN3762] and through recruiting interventions that target earlier stages in the enrollment funnel. Using nearly the same sample of 15 public research universities used in this study, @RN4758 found that 12 universities made more out-of-state visits than in-state visits and visits to out-of-state public high schools focused on affluent, predominantly white schools. Public research universities have a financial incentive to target private school students, who tend to come from affluent households that are able to pay nonresident tuition price without need-based aid. This study extends the literature on nonresident students by analyzing the scale of visits to private schools, the overlap and divergence in recruiting networks between public and private universities, and the characteristics of visited private schools.

We also engage with scholarship on linkages between private schools and private universities, a linkage that sociologists have observed for many decades. Private school students comprise a disproportionate share of enrollment at selective private universities (Author calculations based on HSLS:2009). Practically, these linkages exist because private schools want to send their students to "good" universities while selective private universities want to enroll students who can pay private university tuition and who have desired academic profiles and/or extracurricular activities [@RN3519]. Prior research has often focused on linkages between elite private schools and elite universities [@RN4407; @RN4935; @RN4640]. Whereas elite universities (e.g., Princeton, Stanford) experience strong student demand based on reputation, merely selective private universities must incite demand by expending resources on recruiting. Prior research has not systematically analyzed connections between private schools and private universities that are selective, but not elite. Therefore, this study extends the literature by investigating private school visits for a sample of 14 selective private universities that exhibit variation in geographic location, religious affiliation, and ranking.

@@

Our analyses shed light on the nature of this competition. When making decisions about which private schools to visit, public research universities seem to be mimicking the behavior of particular private universities rather than mimicking the behavior of private universities writ large. Several public universities seem to be competing directly for students targeted by a particular private university. For example, UC-Berkeley visited 27.9% of the schools visited by Northwestern, 28.9% of the schools visited by Tufts, and 27.8% of the schools visited by Emory (Table \@ref(tab:privu-percent-visited-by-pubu-matrix)). We reason UC-Berkeley is competing directly for students because all four universities are in the same broad echelon of prestige. Other potential examples of direct competition include the University of Georgia targeting schools visited by Emory, SMU, TCU, and Tulane, or the University of Arkansas targeting schools visited by Baylor.

However, public universities that made the greatest number of private school visits seemed to be targeting the lower-achieving students at private schools targeted by particular private universities. For example, the University of Alabama visited 52.2% of the 625 private schools that Notre Dame visited, with most of the overlap at Catholic schools. Given the differences in academic prestige between them, the University of Alabama may not be competing for the same students as Notre Dame. Instead, they seem to be targeting Catholic school students of a similar economic echelon, who have lower academic achievements, and want to attend a football powerhouse. Additionally, CU-Boulder visited 42.4% of schools visited by Northwestern, which could be interpreted as CU-Boulder competing for lower-achieving students of a similar economic and cultural echelon at the schools targeted by Northwestern. Similarly, the University of South Carolina competes heavily at schools targeted by Emory, SMU, TCU, and Tulane.

@@

Market research suggests that off-campus recruiting visits will remain a vital means of connecting to students and guidance counselors, just as they were prior to the Covid pandemic. @RN4664 found that "travel" was the top-ranked marketing/recruiting expenditure item for both public universities and private universities. During the height of Covid, off-campus recruiting visits ceased and were partially replaced by virtual counselor events [@parents_covid_rnl]. Admissions counselors are now traveling again. @RN4896 found that travel was the fifth-ranked expenditure item for public universities and the second-ranked expenditure item for private universities. When asked about areas of future expenditure growth, travel was the top-ranked item for public universities and the second-ranked item for private universities [@RN4896], suggesting that universities are increasing travel following the decline of the pandemic.

The @eab_nacac_counselor_2023 survey of high school counselors asked, "How do you prefer to receive information from colleges?" High school visits by college representatives was by far the highest rated "channel" for private school counselors and was also the highest rated channel for public school counselors. By contrast, virtual visits by admissions representatives received low ratings by high school guidance counselors. Finally, @RN4895 asked universities to rank the most effective "outreach strategies for high school counselors." "Calling after visits" was deemed effective for 97% of private universities (the top strategy) and was deemed effective by 88% of public universities. What insights do we glean from these market research findings? In the era of mass marketing and data science, the pursuit of traditional college students from privileged schools and communities is still a hand-to-hand business, and off-campus recruiting visits remain a vital means of recruiting these students.

Considering the practical importance of off-campus recruiting visits, what are the practical implications of our research? Prior scholarly and market research shows that students are sensitive to recruiting interventions, including recruiting visits [@RN4324; @RN4402]. Consequently, the racial and socioeconomic composition of the schools that a university chooses to visit affects the racial and socioeconomic composition of the university. Our research on visits to public high schools [@RN4758] and this study on visits to private schools find that our sample of public research universities and selective private universities tend to focus their recruiting efforts on affluent and predominantly white schools and communities. Some universities may publicly proclaim a commitment to social mobility and privately pursue an intentional recruiting strategy that targets affluent schools and communities. We offer practical guidance for universities that may be unintentionally implementing a recruiting strategy that privileges the privileged.

Several studies suggest that universities -- particularly elite private universities -- pursue racial diversity by recruiting non-white students who attend predominantly white private high schools [e.g., @RN4495]. Additionally, scholarship that engages with enrollment in private schools by Black students suggests that many high-income Black families began sending their children to elite, predominantly white private schools once these schools were willing to enroll Black students [@RN4724]. Therefore, universities may visit predominantly white private high schools in search of non-white students. One concern about this practice is that non-white students at elite private schools may be in high demand by many universities, but similarly achieving non-white students at predominantly non-white public schools might be ignored.

In the aftermath of the Supreme Court decision on race-conscious admissions, @RN4954 published a report titled *Strategies for Increasing Diversity and Opportunity in Higher Education.*  The report recommended that universities diversify the set of high schools they recruit from:

>> Rather than recruiting primarily from private and public high schools composed of predominantly high-income students, institutions can expand and prioritize their outreach to high schools with substantial populations of low-income students and students of color. As part of expanding outreach, institutions can assign admissions recruiters to high schools from which they have not typically recruited to provide equal opportunity for all students.

@@

kable(univ_sample, caption = "Characteristics of universities in analysis sample", booktabs = T, escape = F, align = c('l', 'l', rep('r', 13))) %>%
  collapse_rows(columns = 1, latex_hline = 'major', valign = 'middle') %>%
  column_spec(c(2), width = '4.5cm', latex_valign = 'm') %>% 
  column_spec(c(3), width = '1.8cm', latex_valign = 'm') %>% 
  column_spec(c(4, 5), width = '3.8cm', latex_valign = 'm') %>% 
  column_spec(c(6, 8, 10:14), width = '2.4cm', latex_valign = 'm') %>% 
  column_spec(c(7), width = '2.7cm', latex_valign = 'm') %>% 
  column_spec(c(9, 15), width = '3cm', latex_valign = 'm') %>% 
  # row_spec(0, align = 'c') %>%
  kable_styling(font_size = 13, latex_options = c('scale_down', 'hold_position'))
```

@@

# 49 public research
# pop_pubu <- c('151351', '171100', '180461', '188030', '199193', '147703', '204857', '207388', '214777', '186380', '149222', '196088', '196097', '204796', '100751', '197036', '164155', '104179', '106397', '110635', '110644', '110653', '110662', '110671', '110680', '201885', '126614', '129020', '139959', '141574', '225511', '142285', '155317', '161253', '163286', '166629', '170976', '181464', '187985', '199120', '215293', '217484', '218663', '230764', '234076', '236948', '230728', '234030', '172644')

# 49 private national
# pop_privu <- c('194824', '139658', '190150', '223232', '212054', '193900', '166027', '144050', '130794', '243744', '166683', '198419', '215062', '162928', '182670', '110404', '147767', '217156', '190415', '227757', '152080', '186131', '221999', '131496', '164988', '160755', '213543', '135726', '216597', '228246', '191241', '186867', '131159', '228875', '239105', '127060', '122436', '207971', '179159', '168148', '164924', '195030', '179867', '201645', '197708', '199847', '211440', '121150', '123961')

library(readxl)
cc2000_edition_data <- read_excel(file.path(data_dir, 'cc2000_edition_data.xls'), sheet = 'Data')
pop_pubu <- (cc2000_edition_data %>% filter(`Control (text)` == 'Public', `CC2000 (text)` == 'Doctoral/Research Universities—Extensive'))$UNITID %>% as.character()

usnews_rankings <- read_csv(file.path(data_dir, 'usnews_rankings.csv'), col_select = c('source', 'control', 'rank', 'univ_id'))
pop_privu <- (usnews_rankings %>% filter(source == 'national-universities', control == 'private', rank <= 100))$univ_id %>% as.character()

@@

kable(univ_pop_df, caption = "Characteristics of universities in analysis sample compared to data collection sample", booktabs = T, escape = F, align = 'r', linesep = '') %>%
  add_header_above(header = c(' ' = 1, 'Analysis sample (N=15)' = 3, 'Data collection sample (N=102)' = 3, 'Analysis sample (N=14)' = 3, 'Data collection sample (N=57)' = 3)) %>%
  add_header_above(header = c(' ' = 1, 'Public research universities' = 6, 'Selective private universities' = 6)) %>%
  kable_styling(font_size = 13, latex_options = 'scale_down') 
```

@@

kable(table_1mode_pubu$onemode_pct_df, caption = 'One-mode public university percent matrix', booktabs = T, align = 'r', linesep = '') %>%
  column_spec(c(7, 9, 15), width = '5em') %>%
  column_spec(c(4, 12), width = '5.5em') %>%
  column_spec(c(6, 10, 16), width = '6em') %>%
  column_spec(c(2, 11, 13), width = '6.5em') %>%
  column_spec(c(3, 8), width = '7em') %>%
  column_spec(c(5), width = '7.5em') %>%
  column_spec(c(14), width = '8.8em') %>%
  row_spec(0, align = 'c') %>% 
  kable_styling(font_size = 13, latex_options = 'scale_down') 
```

@@

kable(table_1mode_privu$onemode_pct_df, caption = 'One-mode private university percent matrix', booktabs = T, align = 'r', linesep = '') %>%
  column_spec(c(2, 3, 7:9, 11), width = '7em') %>% 
  column_spec(c(4:6, 10, 12, 13), width = '6em') %>% 
  column_spec(14:15, width = '8.5em') %>%
  row_spec(0, align = 'c') %>% 
  kable_styling(font_size = 13, latex_options = 'scale_down') 
```

@@

kable(table_1mode_u$onemode_pct_df[c(1,7,9,13,16,17,18,21,22,23,24,25,27,28,29),c(2,3,4,5,6,8,10,11,12,14,15,19,20,26)], caption = 'For private universities, the percent of high schools visited by public universities', booktabs = T, align = 'r', linesep = '') %>%
  column_spec(c(2, 3, 7:9, 11), width = '7em') %>% 
  column_spec(c(4:6, 10, 12, 13), width = '6em') %>% 
  column_spec(14:15, width = '8.5em') %>%
  row_spec(0, align = 'c') %>% 
  kable_styling(font_size = 13, latex_options = 'scale_down')
```

@@

kable(table_1mode_u$onemode_pct_df[c(2,3,4,5,6,8,10,11,12,14,15,19,20,26),c(1,7,9,13,16,17,18,21,22,23,24,25,27,28,29)], caption = 'For public universities, the percent of high schools visited by private universities', booktabs = T, align = 'r', linesep = '') %>%
  column_spec(c(7, 9, 15), width = '5em') %>%
  column_spec(c(4, 12), width = '5.5em') %>%
  column_spec(c(6, 10, 16), width = '6em') %>%
  column_spec(c(2, 11, 13), width = '6.5em') %>%
  column_spec(c(3, 8), width = '7em') %>%
  column_spec(c(5), width = '7.5em') %>%
  column_spec(c(14), width = '8.8em') %>%
  row_spec(0, align = 'c') %>% 
  kable_styling(font_size = 13, latex_options = 'scale_down')
```

@@

kable(table_1mode_pubu$onemode_df, caption = 'One-mode public university count matrix', booktabs = T, align = 'r', linesep = '') %>%
  column_spec(c(7, 9, 15), width = '5em') %>%
  column_spec(c(4, 12), width = '5.5em') %>%
  column_spec(c(6, 10, 16), width = '6em') %>%
  column_spec(c(2, 11, 13), width = '6.5em') %>%
  column_spec(c(3, 8), width = '7em') %>%
  column_spec(c(5), width = '7.5em') %>%
  column_spec(c(14), width = '8.8em') %>%
  row_spec(0, align = 'c') %>% 
  kable_styling(font_size = 13, latex_options = 'scale_down') 
```

@@

kable(table_1mode_privu$onemode_df, caption = 'One-mode private universities count matrix', booktabs = T, align = 'r', linesep = '') %>%
  column_spec(c(2, 3, 7:9, 11), width = '7em') %>% 
  column_spec(c(4:6, 10, 12, 13), width = '6em') %>% 
  column_spec(14:15, width = '8.5em') %>%
  row_spec(0, align = 'c') %>% 
  kable_styling(font_size = 13, latex_options = 'scale_down') 
```

### 2023-12-15 r_r/tables_and_figures.Rmd
updates to manuscript

kable(univ_pop_df, caption = "Characteristics of universities in analysis sample compared to data collection sample", booktabs = T, escape = F, align = 'c', linesep = '') %>%
  add_header_above(header = c(' ' = 1, 'Analysis sample (N=15)' = 3, 'Data collection sample (N=49)' = 3, 'Analysis sample (N=14)' = 3, 'Data collection sample (N=49)' = 3)) %>%
  add_header_above(header = c(' ' = 1, 'Public research universities' = 6, 'Selective private universities' = 6)) %>%
  kable_styling(font_size = 13, latex_options = 'scale_down') 
```

>>>>

kable(univ_pop_df, caption = "Characteristics of universities in analysis sample compared to data collection sample", booktabs = T, escape = F, align = 'c', linesep = '') %>%
  add_header_above(header = c(' ' = 1, 'Analysis sample (N=15)' = 3, 'Data collection sample (N=102)' = 3, 'Analysis sample (N=14)' = 3, 'Data collection sample (N=57)' = 3)) %>%
  add_header_above(header = c(' ' = 1, 'Public research universities' = 6, 'Selective private universities' = 6)) %>%
  kable_styling(font_size = 13, latex_options = 'scale_down') 
```
